---
title: "RhizMicrobiome_GenusLevel"
author: "Sara Colom"
date: "2/8/2020"
output: 
  github_document:
    toc: true
---


## Sample sizes

### Table sample number by species and treatment


| Species          | Treatment       | N      |
|--------------    |-------------    |----    |
| I. purpurea      | Alone           | 27     |
| I. purpurea      | Competition     | 78     |
| I. hederacea     | Competition     | 78     |


### Table number of maternal line per species

| Species          | Number of ML     |
|--------------    |--------------    |
| I. purpurea      |      10          |
| I. hederacea     |       5          |



# Load Libraries

```{r load_libs, include=T,warning=F,comment=F,message=F}
library(phyloseq)
library(ggplot2)
library(ape)
library(vegan)
library(plyr)
library(dplyr)
library(scales)
library(grid)
library(reshape2)
library(pegas)
library(pgirmess)
library(multcomp)
library(multcompView)
library(ggpubr)
library(ggcorrplot)
library(RColorBrewer)
library(plotly)
library(ggthemes)
library(corrplot)
library(Hmisc)
library(emmeans)
library(lmerTest)


source("miSeq.R")

# Aesthetics
Tx<-theme(axis.text.y = element_text(size = 20),
          axis.title.y = element_text(size = 20)) +
          theme(axis.text.x = element_text(vjust = 1, hjust=1, angle=0, size = 20),
          axis.title.x = element_text(angle=0, size = 20),
          plot.title=element_text(size = 25,hjust=0))

# Margins
Margin<-theme(
  panel.background = element_rect(fill = "white"),
  plot.margin = margin(2, 2, 2, 2, "cm"),
  plot.background = element_rect(
    fill = "white",
    colour = "black",
    size = 1
  )
)

# Aesthetics
Tx2<-theme(axis.text.y = element_text(size = 20),
          axis.title.y = element_text(size = 25)) +
          theme(axis.text.x = element_text(vjust = 1, hjust=1, size = 25),
          axis.title.x = element_text(size = 25),
          plot.title=element_text(size = 25,hjust=0))

GoldGrey=c("#F1CE63", "#79706E")
GreenBlue=c("#59A14F", "#4E79A7")

```



# Read in Data

```{r Import Data,warning=F,comment=F,message=F}
### loading mothur output with FWDB+silva taxonomy and sample metadata. 
### Experiments run in 


sharedfile <- "../DataSets/stability.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.opti_mcc.shared"
taxfile <- "../DataSets/stability.trim.contigs.good.unique.good.filter.unique.precluster.pick.pick.opti_mcc.0.03.cons.taxonomy"

mothurdata <- import_mothur(mothur_shared_file = sharedfile, mothur_constaxonomy_file = taxfile)

sampledata <- read.csv('../DataSets/MetaDataTest.csv')

SAMPLE <- sampledata
row.names(SAMPLE)=SAMPLE$Sample_ID

SAMPLE <- subset(SAMPLE, SAMPLE$TRT == "Alone"|SAMPLE$TRT == "Inter")
SAMPLE <- sample_data(SAMPLE)

### create phyloseq object
physeq.all <- merge_phyloseq(mothurdata, SAMPLE) # Modified version worked


### We need to change the taxonomy names: when using the fwdb taxonomy we need to add different headers after removing the last column, which contains no information except for 1 Verrucomicrobia taxon

#   tax_table(physeq.all) <- tax_table(physeq.all)[,-7] # this removes the final column
colnames(tax_table(physeq.all))<-c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus")


#   tax_table(physeq.all) <-cbind(tax_table(physeq.all),row.names(tax_table(physeq.all)))

### removing non-bacterial reads (was already done in mothur, but just to be safe after merging taxonomies)
#   physeq.all <- subset_taxa(physeq.all, Kingdom == "Bacteria")

```



```{r, aggregate data by genus}

## Aggregate at the genus level, then saving ea componant externally to reload and save time.

        # physeq_genus <- physeq.all %>%
         #tax_glom(taxrank = "Genus") 

        # otu=data.frame(otu_table(physeq_genus))
        
        # meta=data.frame(sample_data(physeq_genus))
        # tax=data.frame(tax_table(physeq_genus))
        
        # write.csv(x = otu,file = "otu_table.csv",row.names = TRUE)
        # write.csv(x = tax,file = "tax_table.csv")# 
        # write.csv(x = meta,file = "sample_table.csv",row.names = F)


otu <- read.csv("../DataSets/otu_table.csv", row.names = 1)

colnames(otu) <- gsub('X', "", colnames(otu))

otu <- otu_table(otu, taxa_are_rows = TRUE)


tax <- read.csv('../DataSets/tax_table.csv', row.names = 1)
taxRows <- row.names(tax)
taxCols <- colnames(tax)
tax <- tax_table(as.matrix(tax))

#row.names(tax)=taxRows
#colnames(tax)=taxCols


meta <- read.csv('../DataSets/MetaDataTest.csv')
namesKeep <- colnames(otu)

row.names(meta) <- meta$Sample_ID

meta <- meta[which(meta$Sample_ID %in% namesKeep), ]

meta <- sample_data(meta)
   
physeq_genus <- merge_phyloseq(otu, tax, meta)

```



```{r Proteobacteria phylum change to class, warning=F,comment=F,message=F}


### ADD THE PROTEOBACTERIA CLASSES TO THE PHYLA NAME FIELD IN PHYLOSEQ OBJECT TAXONOMY 

phy <- data.frame(tax_table(physeq_genus))
Phylum <- as.character(phy$Phylum)
Class <- as.character(phy$Class)
for  (i in 1:length(Phylum)){ 
  if (Phylum[i] == "Proteobacteria"){
    if (Class[i] == "unclassified"){
      Phylum[i] <- Phylum[i]       
    } else {
      Phylum[i] <- Class[i]
    }
  } 
}


Phylum <- as.character(phy$Phylum)
Class <- as.character(phy$Class)
phy$Phylum <- Phylum
t <- tax_table(as.matrix(phy))

```


```{r read in phylum data, include=FALSE,warning=F,comment=F,message=F}
SAMPLE <- sample_data(physeq_genus)

SAMPLE$Species <- ifelse(grepl("Ip", SAMPLE$ML), "Ip", "Ihed")

physeq.all2 <- merge_phyloseq(physeq_genus, SAMPLE) 
```

## Sequence depth and pruning

```{r data scaling}
physeq <- (physeq.all2)
physeq <- prune_taxa(taxa_sums(physeq) > 0, physeq) # Remove taxa with no counts

#check number of reads in each sample, differences in count are in part due differet numbers of chlorophyl reads depending on time of experiment

# Make a data frame with a column for the read counts of each sample
sample_sum_df <- data.frame(sum = sample_sums(physeq))

# Histogram of sample read counts
ggplot(sample_sum_df, aes(x = sum)) + 
  geom_histogram(color = "black", fill = "#59A14F", binwidth = 2500) +
  ggtitle("Distribution of sample sequencing depth") + 
  xlab("Read counts") +
  theme(axis.title.y = element_blank()) +
  theme_classic()


# Scales reads to smallest library size 
source("https://raw.githubusercontent.com/michberr/MicrobeMiseq/master/R/miseqR.R")
#physeq.scale <- scale_reads(physeq, min(sample_sums(physeq)))

##### Normalization #######

# Scales reads by 
# 1) taking proportions,
# 2) multiplying by a given library size of n
# 3) rounding down

physeq1 <- prune_samples(sample_sums(physeq) > 20000, physeq)

sample_sum_df1 <- data.frame(sum = sample_sums(physeq1))

# Histogram of sample read counts
ggplot(sample_sum_df1, aes(x = sum)) + 
  geom_histogram(color = "black", fill = "#59A14F", binwidth = 2500) +
  ggtitle("Distribution of sample sequencing depth after pruning") + 
  xlab("Read counts") +
  theme(axis.title.y = element_blank()) +
  theme_classic()


n <- min(sample_sums(physeq1))
  physeq.scale <-
    transform_sample_counts(physeq1, function(x) {
      (n * x/sum(x))  # Transform to rel. abundance
    })
  
physeq.scale <- prune_taxa(taxa_sums(physeq.scale) > 0, physeq.scale)


physeq1

```


```{r}
physeq.bray <- phyloseq::distance(physeq = physeq.scale, method = "bray")

# # # # # # # # # # # # # # # # # # # # # 
# Subsample for within I. purpurea only
# # # # # # # # # # # # # # # # # # # # # 

physeq.Purp <- subset_samples(physeq.scale, Species == "Ip")
sampledf.Purp<- data.frame(sample_data(physeq.Purp))

# Calculate bray curtis for summer samples only
physeq.Purp.bray <- phyloseq::distance(physeq = physeq.Purp, method = "bray")

```

## Test for differently abundant OTUs

### Visualize the different groups at the family level
```{r}
# Check for 'core' microbiome members as the phylum level which taxa are present
# at 1 % overall abundance and at least 75% of samples 

PhylumGlom <- tax_glom(physeq.Purp,taxrank = "Phylum")
FamGlom <- tax_glom(physeq.Purp,taxrank = "Family")
ClassGlom <- tax_glom(physeq.Purp,taxrank = "Class")


coreTaxa <- filter_taxa(PhylumGlom, function(x) sum(x > 1) > (0.75 * length(x)), TRUE)

coreTaxaFamily <- filter_taxa(FamGlom, function(x) sum(x > 1) > (0.75 * length(x)), TRUE)

## Plot at phylum level


physeq1_phylumAlone <- coreTaxa  %>%
  subset_samples(TRT == "Alone")%>%
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt()      


physeq1_phylumComp <- coreTaxa  %>%
  subset_samples(TRT != "Alone")%>%
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt()       

physeq1_phylum <- rbind(physeq1_phylumAlone,physeq1_phylumComp)

# Melt to long format
physeq1_phylum <- physeq1_phylum[order(physeq1_phylum$Phylum), ] 

#colnames(physeq1_phylum)[13]="Phylum"

# Sort data frame alphabetically by phylum
# Set colors for plotting
phylum_colors <- c(
  "#CBD588", "#5F7FC7", "orange", "#DA5724", "#508578", "#CD9BCD",
   "#AD6F3B", "#673770", "#D14285", "#652926", "#C84248", 
  "#8569D5", "#5E738F", "#D1A33D", "#8A7C64", "#599861"
)


#colnames(physeq1_phylum)[which(names(physeq1_phylum)%in%"Rank2")]="Phylum"

# Plot 
ggplot(physeq1_phylum, aes(x = TRT, y = Abundance, fill = Phylum)) + 
  geom_bar(stat = "identity",position='dodge') +
  #scale_fill_manual(values = phylum_colors) +
  theme(axis.title.x = element_blank()) + 
  ylab("Relative Abundance (Phyla > 75 %) \n") +
  ggtitle("Phylum Composition  \n Bacterial Communities by Sampling per Treatment") +
  theme_classic() +
  ylab("Relative Abundance")

```


## Testing if different at the phylum level
```{r}

PhlyaList <- as.list(unique(physeq1_phylum$Phylum))

KrusMicrTest <- function(Microbe){
  Result = kruskal.test(Abundance ~ TRT, data = physeq1_phylum[which(physeq1_phylum$Phylum == Microbe),]) 

ResDf <- data.frame("ChiSq" = Result$statistic, "Pval"=Result$p.value, "Phylum" = paste(Microbe))  
  return(ResDf)  
}

AllResults <- lapply(PhlyaList, KrusMicrTest)
KruskalRes <- do.call('rbind', AllResults)
```


## Family level
```{r}

# Family level
physeq1_FamilyAlone <- coreTaxaFamily  %>%
  subset_samples(TRT == "Alone")%>%
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt()                                      # Melt to long format

physeq1_FamilyComp <- coreTaxaFamily  %>%
  subset_samples(TRT != "Alone")%>%
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt()       

physeq1_Family <- rbind(physeq1_FamilyAlone,physeq1_FamilyComp)


FamList <- as.list(unique(physeq1_Family$Family))

KrusMicrTest<-function(Microbe){
  Result <- kruskal.test(Abundance ~ TRT, data = physeq1_Family[which(physeq1_Family$Family == Microbe), ]) 

ResDf <- data.frame("ChiSq" = Result$statistic, "Pval"= Result$p.value, "Family" = paste(Microbe))  
  return(ResDf)  
}

AllResultsFamily <- lapply(FamList, KrusMicrTest)
KruskalResFamily <- do.call('rbind', AllResultsFamily)

KruskalResFamily %>% 
  filter(Pval < 0.055)

# Do false discovery rate for mulitple corrections

KruskalResFamily$AdjPval <- p.adjust(KruskalResFamily$Pval, method = "fdr", n = 213)
```

## Alpha Diversity

```{r alpha diversity,warning=F,comment=F,message=F}
hist(sample_sums(physeq1))

# Remove sample with less than 20K reads, this looks WAAAY off
# Rarify first
ps.rarefied <- rarefy_even_depth(physeq1, rngseed = 1, sample.size = min(sample_sums(physeq1)), replace = F)

plot_richness(ps.rarefied,x = "TRT",measures = c("Observed", "Shannon")) +
  geom_boxplot() +
  theme_classic()

```


## Alpha Diversity

#### Note: Measures of alpha-diversity 

_*Inverse Simpson*_ it is an indication of the richness in a community with uniform evenness that would have the same level of diversity. So while measures such as the Shannon index are somewhat abstract, the inverse of the Simpson index has some biological interpretation. Other advantages of the Simpson-based metrics are that they do not tend to be as affected by sampling effort as the Shannon index.
   
- _*Species richness*_ is simply a count of species, and it does not take into account the abundances of the species or their relative abundance distributions.
   
- _*Simpson's Diversity Index*_ is a measure of diversity which takes into account the number of species present, as well as the relative abundance of each species. As species richness and evenness increase, so diversity increases.



```{r Richness and Inverse Simpson, warning = F, comment = F, message = F}

### I commented out the subsampling to save time. The matrices produced were saved and then reoponed from their tab delmited formats into corresponding matrices.

set.seed(711)

r <- rarefy_even_depth(physeq1, sample.size = n, verbose = FALSE, replace = TRUE, rngseed = 711)
         
 ## Calculate richness
rich <- as.numeric(as.matrix(estimate_richness(r, measures = "Observed")))

## Calculate Inverse Simpson

simp <- as.numeric(as.matrix(estimate_richness(r, measures = "InvSimpson")))

sim<- as.numeric(as.matrix(estimate_richness(r, measures = "Simpson")))

s <- rarefy_even_depth(physeq1, sample.size = n, verbose = FALSE, replace = TRUE, rngseed = 711)

shan <- as.numeric(as.matrix(estimate_richness(r, measures = "Shannon")))

InvSimp <- simp

```



```{r}
# Create a new dataframe to hold the means and standard deviations of richness estimates

length(rich)

Sample_ID <- sample_names(physeq1)
Block <- sample_data(physeq1)$Block
Species <-sample_data(physeq1)$Species
TRT <- sample_data(physeq1)$TRT
Combos <-sample_data(physeq1)$Combos
ML <- sample_data(physeq1)$ML

alpha <- data.frame(Sample_ID,ML, Block, TRT, Species, Combos, rich, InvSimp, sim, shan)

alpha$even <- alpha$shan/alpha$rich

# DO VIOLIN PLOT HERE!!!


p <- ggplot(alpha %>% filter(TRT == "Inter"), aes(x = Species, y = rich)) +
  geom_violin(trim = FALSE, aes(fill = Species), alpha = 0.3) + 
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 1, aes(color = Species, fill = Species)) +
  theme_classic() +
  scale_color_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Mean Species Richness")


q <- ggplot(alpha %>% filter(TRT == "Inter"), aes(x = Species, y = InvSimp)) +
  geom_violin(trim = FALSE, aes(fill = Species), alpha = 0.3) + 
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 1, aes(color = Species, fill = Species)) +
  theme_classic() +
  scale_color_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Mean Species Inverse Simpson") +
  ylab("")


t <- ggplot(alpha %>% filter(TRT == "Inter"), aes(x = Species, y = sim)) +
  geom_violin(trim = FALSE, aes(fill=Species), alpha = 0.3) + 
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 1, aes(color = Species, fill = Species)) +
  theme_classic() +
  scale_color_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Mean Simpson") +
  ylab("")

v <- ggplot(alpha %>% filter(TRT == "Inter"), aes(x = Species, y = even)) +
  geom_violin(trim = FALSE, aes(fill = Species), alpha = 0.3) + 
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 1, aes(color = Species, fill = Species)) +
  theme_classic() +
  scale_color_brewer(palette = "Paired") +
  scale_fill_brewer(palette = "Paired") +
  ggtitle("Mean Evenness") +
  ylab("")



ggarrange(p, q, t, v, common.legend = T, ncol = 2, nrow = 2)



      ### Treatment

pT <- ggplot(alpha %>% filter(Species == "Ip"), aes(x = TRT, y = rich)) +
  geom_violin(trim = FALSE, aes(fill = TRT), alpha = 0.3) + 
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 1, aes(color = TRT,fill = TRT)) +
  theme_classic() +
  scale_color_manual(values = c("#B2DF8A", "#fee090")) +
  scale_fill_manual(values = c("#B2DF8A", "#fee090")) +
  ggtitle("Mean Treatment Richness")


qT <- ggplot(alpha %>% filter(Species == "Ip"), aes(x = TRT, y = InvSimp)) +
  geom_violin(trim = FALSE, aes(fill = TRT), alpha = 0.3) + 
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize=1, aes(color = TRT, fill = TRT)) +
  theme_classic() +
  scale_color_manual(values = c("#B2DF8A", "#fee090")) +
  scale_fill_manual(values = c("#B2DF8A", "#fee090")) +
  ggtitle("Mean Treatment Inverse Simpson") +
  ylab(" ")

ggarrange(pT, qT, common.legend = T)

```

## Test for differences
# Alpha Diversity distribution

```{r, warning=F,comment=F,message=F}
# first check for normality. To test for normalcy statistically, we can run the Shapiro-Wilk test of normality.

shapiro.test(alpha$rich) # Not normal
shapiro.test(alpha$even) # Normal
shapiro.test(alpha$sim) # Normal

histogram(alpha$rich,xlab = "Richness") # Note richness though, not normal, looks close to normal.

histogram(alpha$InvSimp, xlab = "Inverse-Simpson")
histogram(alpha$sim, xlab = "Simpson")
```


# Linear mixed models

```{r, LMM within species Not Reported}
# Test for treatment within I. purpurea


RichLmm <- lmer(rich ~ TRT + Block + (1|Block:ML), alpha %>% filter(Species == "Ip"))

InvLmm <- lmer(InvSimp ~ TRT + Block + (1|Block:ML), alpha %>% filter(Species == "Ip"))

SimLmm <- lmer(sim ~ TRT + Block + (1|Block:ML), alpha %>% filter(Species == "Ip"))

EvenLmm <- lmer(even ~ TRT + Block + (1|Block:ML), alpha %>% filter(Species == "Ip"))

anova(RichLmm)
ranova(RichLmm)

anova(InvLmm)
ranova(InvLmm)

anova(SimLmm)
ranova(SimLmm)

anova(EvenLmm)
ranova(EvenLmm)
```


# ANOVA Test for treatment within I. purpurea (Table 1)

```{r ANOVA Test for treatment within I. purpurea, warning=F,comment=F,message=F}

alpha.purp <- alpha %>% filter(Species == "Ip")

aov.Richness <- lm(rich ~ TRT + Block, alpha.purp)
aov.simpsonInv <- lm(InvSimp ~ TRT + Block, alpha.purp)
aov.simpson <- lm(sim ~ TRT + Block, alpha.purp)
aov.evenness <- lm(even ~ TRT + Block, alpha.purp)
aov.shannon <- lm(shan ~ TRT + Block, alpha.purp)


#Call for the summary of that ANOVA, which will include P-values
anova(aov.Richness)
anova(aov.simpsonInv)
anova(aov.simpson)
anova(aov.evenness)

# No significant differences
```




# Community Composition

## Beta Diversity


```{r PCoA Treatment and Species comparisoin Bray Curtis Estimate, warning = F, comment = F, message = F, fig.height = 12, fig.width = 16}

# Beta diversity pcoa Bray-Curtis DNA only 
physeq.pcoa <-
  ordinate(
    physeq = physeq.scale,
    method = "PCoA",
    distance = "bray"
  )

physeq.pcoa.vectors <- data.frame(physeq.pcoa$vectors[, 1:4])

physeq.pcoa.vectors$Duplicates <- row.names(physeq.pcoa.vectors)

SampData <- data.frame(sample_data(physeq))

colnames(SampData)[1] <- "Duplicates"

SampData <- subset(SampData, SampData$TRT == "Inter"|SampData$TRT == "Alone")

physeq.pcoa.df <- droplevels(merge(physeq.pcoa.vectors, SampData,by="Duplicates"))

bray_values <- physeq.pcoa$values
bray_rel_eigens <- bray_values$Relative_eig

bray_rel_eigen1 <- bray_rel_eigens[1]
bray_rel_eigen1_percent <- round(bray_rel_eigen1 * 100, digits = 1)

bray_rel_eigen2 <- bray_rel_eigens[2]
bray_rel_eigen2_percent <- round(bray_rel_eigen2 * 100, digits = 1)

bray_rel_eigen3 <- bray_rel_eigens[3]
bray_rel_eigen3_percent <- round(bray_rel_eigen3 * 100, digits = 1)

bray_rel_eigen4 <- bray_rel_eigens[4]
bray_rel_eigen4_percent <- round(bray_rel_eigen4 * 100, digits = 1)

bray_rel_eigen5 <- bray_rel_eigens[5]
bray_rel_eigen5_percent <- round(bray_rel_eigen5 * 100, digits = 1)

# Prep axis labels & title
bray_axis1 <- paste("PCoA 1:",bray_rel_eigen1_percent, "%")
bray_axis2 <- paste("PCoA 2:",bray_rel_eigen2_percent, "%")
bray_axis3 <- paste("PCoA 3:",bray_rel_eigen3_percent, "%")
bray_axis4 <- paste("PCoA 4:",bray_rel_eigen4_percent, "%")
PCoA_title <- paste("Bray-Curtis, ",ntaxa(physeq.scale), "OTUs")


pcoa_exp_trt <- ggplot(physeq.pcoa.df, aes(Axis.1, Axis.2, color = TRT, fill = TRT)) +
  xlab(bray_axis1) + 
  ylab(bray_axis2) +
  geom_point(alpha = 0.9) + 
  theme_classic() +
  scale_fill_manual(values = c("#00B050", "grey"), "Treatment", labels = c("Alone", "Competition")) +
  scale_color_manual(values = c("#00B050", "grey"), "Treatment", labels = c("Alone", "Competition")) +
  stat_ellipse(geom = "polygon", alpha = 1/6, aes(fill = TRT)) +
  Tx +
  theme(axis.text.x = element_text(angle = 45), axis.text = element_text(color="black", size = 15,vjust = 0.5,hjust = 1))
#  scale_fill_discrete(name = "Treatment", labels = c("Alone", "Competition")) +
 # scale_color_discrete(name = "Treatment", labels = c("Alone", "Competition"))



pcoa_exp_trtA <- ggplot(physeq.pcoa.df, aes(Axis.3, Axis.4, color = TRT,fill = TRT)) +
  xlab(bray_axis3) + 
  ylab(bray_axis4) + 
  geom_point(alpha = 0.9) + 
  theme_classic() +
  scale_fill_manual(values = GoldGrey, "Treatment", labels = c("Alone", "Competition")) +
  scale_color_manual(values = GoldGrey, "Treatment", labels = c("Alone", "Competition")) +
  stat_ellipse(geom = "polygon", alpha = 1/6, aes(fill = TRT)) +
  Tx


pcoa_exp_sp <- ggplot(physeq.pcoa.df, aes(Axis.1, Axis.2, color = Species,fill = Species)) +
  xlab("") + 
  ylab(bray_axis2) +
  geom_point(alpha = 0.9) + 
  theme_classic() +
 scale_fill_manual(values = GreenBlue, "Species", labels = c("I. hederacea", "I. purpurea")) +
  scale_color_manual(values = GreenBlue, "Species", labels = c("I. hederacea", "I. purpurea")) +
   stat_ellipse(geom = "polygon", alpha = 1/6, aes(fill = Species)) +
  Tx


pcoa_exp_spA <- ggplot(physeq.pcoa.df, aes(Axis.3, Axis.4, color = Species,fill = Species)) +
  xlab("") + 
  ylab(bray_axis4) +
  geom_point(alpha = 0.9) + 
  theme_classic() +
  scale_fill_manual(values = GreenBlue, "Species", labels = c("I. hederacea", "I. purpurea")) +
  scale_color_manual(values = GreenBlue, "Species", labels = c("I. hederacea", "I. purpurea")) +
  stat_ellipse(geom = "polygon", alpha = 1/6, aes(fill = Species)) +
  Tx

A <- ggarrange(pcoa_exp_trt,pcoa_exp_trtA,common.legend = T, Labels = c("C", "D"),font.label = list(size = 15, color = "black", face =
  "plain"),hjust = -8,vjust = 1)

B <- ggarrange(pcoa_exp_sp,pcoa_exp_spA,common.legend = T, labels = c("A", "B"),font.label = list(size = 15, color = "black", face =
  "plain"),hjust = -8,vjust = 1)

MainFig <- ggarrange(B, A, nrow = 2)

annotate_figure(MainFig,
                top = text_grob(PCoA_title, face = "plain", size = 25, lineheight = 2)
)
```



```{r 3D graphic, warning = F, comment = F, message = F}

physeq.pcoa.df$TrtSp <- paste(physeq.pcoa.df$TRT, physeq.pcoa.df$Species, sep="")

p <- plot_ly(physeq.pcoa.df, x = ~Axis.1, y = ~Axis.2, z = ~Axis.3, color = ~TrtSp) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = bray_axis1),
                     yaxis = list(title = bray_axis2),
                     zaxis = list(title = bray_axis3)))


p1 <- plot_ly(physeq.pcoa.df, x = ~Axis.1, y = ~Axis.2, z = ~Axis.4, color = ~Species) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = bray_axis1),
                     yaxis = list(title = bray_axis2),
                     zaxis = list(title = bray_axis4)))

p3 <- plot_ly(physeq.pcoa.df, x = ~Axis.1, y = ~Axis.2, z = ~Axis.4, color = ~Species) %>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = bray_axis1),
                     yaxis = list(title = bray_axis3),
                     zaxis = list(title = bray_axis4)))

```

# PERMANOVA (Table 2)

```{r adonis experimental data when all data combined, warning = F,comment = F,message = F}

## All experiment, add variable treatment, then species
sampledf <- data.frame(sample_data(physeq.scale))

physeq.purp = subset_samples(physeq.scale, Species == "Ip")

treatment <- sampledf %>% 
  filter(Species == "Ip") %>% 
  pull(TRT)

block <- sampledf %>% 
  filter(Species == "Ip") %>% 
  pull(Block)

adonis(otu_table(physeq.purp) %>% t ~ treatment + block, method = "bray")

```


# Correlations with root traits


## Prep root data

```{r Preliminary exam on root linear model}
RootData <- read.csv("../DataSets/RootTraits_PCs.csv")

RootAlphaObs <- merge(alpha, RootData[c("Sample_ID", "PC1", "PC2", "PC3", "PC4")])
```

## Within species (root traits and alphadiv)


```{r Preliminary exam on roots within species}

library(multcomp)

#summary(glht(SimpInvPC1, mcp(rank="Tukey")))

#################################################################
##################### SUBSET for I.purpurea #####################
#################################################################


RootAlphaPurp <- droplevels(RootAlphaObs %>% filter(Species == "Ip"))
RootAlphaPurp$Comp <- sub(".*\\-", "", RootAlphaPurp$Combos)

### Linear regressions

SimpPC1 <- lm(sim ~ PC1 + Block + TRT, RootAlphaPurp) 
summary(SimpPC1)

SimpInvPC1 <- lm(InvSimp ~ PC1 + Comp + TRT + Block, RootAlphaPurp) 
summary(SimpInvPC1)

RichPC1 <- lm(rich ~ PC1 + Block + TRT, RootAlphaPurp) 
summary(RichPC1)

EvenPC1 <- lm(even ~ PC1 + Block + TRT , RootAlphaPurp) 
summary(EvenPC1)


SimpPC2 <- lm(sim ~ PC2 + Block + TRT, RootAlphaPurp) 
summary(SimpPC2)

SimpInvPC2 <- lm(InvSimp ~ PC2 + Comp + TRT + Block, RootAlphaPurp) 
summary(SimpInvPC2)

RichPC2 <- lm(rich ~ PC2 + Block + TRT, RootAlphaPurp) 
summary(RichPC2)
car::Anova(mod = lm(rich ~ PC2 + Block + TRT*PC2, RootAlphaPurp), type = "III")

EvenPC2 <- lm(even ~ PC2 + Block + TRT , RootAlphaPurp) 
summary(EvenPC2) 

SimpPC3 <- lm(sim ~ PC3 + Block + TRT, RootAlphaPurp) 
summary(SimpPC3)

SimpInvPC3 <- lm(InvSimp ~ PC3 + Comp + TRT + Block, RootAlphaPurp) 
summary(SimpInvPC3)


RichPC3 <- lm(rich ~ PC3 + Block + TRT, RootAlphaPurp) 
summary(RichPC3)


EvenPC3 <- lm(even ~ PC3 + Block + TRT , RootAlphaPurp) 
summary(EvenPC3)


SimpPC4 <- lm(sim ~ PC4 + Block + TRT, RootAlphaPurp) 
summary(SimpPC4)
car::Anova(mod = lm(sim ~ PC4 + Block + TRT*PC4, RootAlphaPurp), type = "III")

# Repeat & remove outlier
SimpPC4_out <- lm(sim ~ PC4 + Block + TRT, RootAlphaPurp %>% filter(sim > 0.94)) 
summary(SimpPC4_out)

SimpInvPC4 <- lm(InvSimp ~ PC4 + Comp + TRT + Block, RootAlphaPurp) 
summary(SimpInvPC4)


RichPC4 <- lm(rich ~ PC4 + Block + TRT, RootAlphaPurp) 
summary(RichPC4)


EvenPC4 <- lm(even ~ PC4 + Block + TRT , RootAlphaPurp) 
summary(EvenPC4)

```


## Plotting significant linear associations 
```{r n}

# Richness and root architecture
P2.rich <- ggplot() +
  geom_point(data = RootAlphaPurp, aes(PC2,rich), alpha = 0.5, size = 5) +
  geom_smooth(data = RootAlphaPurp, method = "lm", aes(PC2,rich), fullrange = TRUE) +
  theme_classic() +
  ylab("Richness") +
  xlab("") +
  Tx2
#+ 
  #annotate("text", x = -1, y = 500, label = "paste(italic(R) ^ 2, \" = 0.15\")", parse = TRUE,hjust=0, size=5) + 
 # annotate("text", x = -1, y = 490, label = "paste(italic(B), \" = -4.65 +/-1.81\")", parse = TRUE,hjust=0, size=5) +
 # annotate("text", x = -1, y = 480, label = "paste(italic(P), \" = 0.01\")", parse = TRUE,hjust=0, size=5)

# Evenness and root architecture
P2.even <- ggplot() +
  geom_point(data = RootAlphaPurp, aes(PC2, even), alpha = 0.5, size = 5) +
  geom_smooth(data = RootAlphaPurp, method = "lm", aes(PC2, even), fullrange = TRUE) +
  theme_classic() +
  ylab("Evenness") +
  xlab("") +
  Tx2#+ 
  #annotate("text", x = -4.5, y = .0114, label = "paste(italic(R) ^ 2, \" = 0.12\")", parse = TRUE,hjust=0, size=5) + 
#  annotate("text", x = -4.5, y = .0112, label = "paste(italic(B), \" = -7.29E-5 +/-3.28E-5\")", parse = TRUE,hjust=0, size=5) +
 # annotate("text", x = -4.5, y =.011, label = "paste(italic(P), \" = 0.04\")", parse = TRUE,hjust=0, size=5)


# Root morphology on species diversity Simpson metric

P4.Sim <- ggplot() +
  geom_point(data = RootAlphaPurp, aes(PC4, sim), alpha = 0.5, size = 5) +
  geom_smooth(data = RootAlphaPurp, method = "lm", aes(PC4, sim), fullrange = TRUE) +
  theme_classic() +
  ylab("Simpson") +
  xlab("") +
  Tx2#+
  #annotate("text", x = -5, y = .988, label = "paste(italic(R) ^ 2, \" = 0.18\")", parse = TRUE,hjust=0, size=5) + 
  #annotate("text", x = -5, y = .985, label = "paste(italic(B), \" = 2.22E-4 +/-6.99E-4\")", parse = TRUE,hjust=0, size=5) +
 # annotate("text", x =-5, y = .982, label = "paste(italic(P), \" < 0.01\")", parse = TRUE,hjust=0, size=5)


P4.simIn <- ggplot() +
  geom_point(data = RootAlphaPurp, aes(PC4, InvSimp), alpha = 0.5, size = 5) +
  geom_smooth(data = RootAlphaPurp, method = "lm", aes(PC4, InvSimp), fullrange = TRUE) +
  theme_classic() +
  ylab("Inverse Simpson") +
  xlab("") +
  Tx2#+
 # annotate("text", x = -5, y = 60, label = "paste(italic(R) ^ 2, \" = 0.12\")", parse = TRUE,hjust=0, size=5) + 
 # annotate("text", x = -5, y = 57, label = "paste(italic(B), \" = 2.08 +/-1.05\")", parse = TRUE,hjust=0, size=5) +
 # annotate("text", x =-5, y = 54, label = "paste(italic(P), \" = 0.02\")", parse = TRUE,hjust=0, size=5)


```


## Linear mixed models (not reported)

```{r}

### Simpson
SimpLMM <- lmer(sim ~ TRT + Block + (1|ML), alpha %>% filter(Species == "Ip")) 
anova(SimpLMM)
ranova(SimpLMM)

### Inverse Simpson
SimpInvLMM <- lmer(InvSimp ~ TRT + Block + (1|ML), alpha %>% filter(Species == "Ip")) 
anova(SimpInvLMM)
ranova(SimpInvLMM)

### Inverse Simpson
RichLMM <- lmer(rich ~ TRT + Block + (1|ML), alpha %>% filter(Species == "Ip")) 
anova(RichLMM)
ranova(RichLMM)

```


```{r Analyzing leaf data, warning=F,comment=F,message=F}
LeafData <- read.csv("../DataSets/SizeFitData.csv")

LeafData$Sample_ID <- paste(LeafData$Position, ifelse(grepl("Ihed", LeafData$ML), "H", "P"), sep="")
```


```{r Read in fitness data}
Fitness = read.csv("../DataSets/FitPA4.csv")

library(dplyr)
# Calculate relative fitness
  # First calculate mean seed number by species and treatment---note* we only have seed output of I. purpurea
MeanSeedNumber <- aggregate(SeedNumber ~ Trt + Species, Fitness, mean)

colnames(MeanSeedNumber) <- c("Trt", "Species", "MeanSeedNumber")

Ipurp.Fit <- Fitness %>%
  filter(Species == "Ip")

Ipurp.Alpha <- alpha %>%
  filter(Species == "Ip")

FitnessPurp <- merge(Ipurp.Fit, MeanSeedNumber, by=c("Trt", "Species"))
FitnessPurp$RelativeFit <- FitnessPurp$SeedNumber/FitnessPurp$MeanSeedNumber
FitnessPurp$Block <- as.factor(FitnessPurp$Block)

FitnessPurp2 <- merge(FitnessPurp, LeafData)
str(FitnessPurp2)
FitnessPurp2$Leaf.Number <- as.numeric(as.character(FitnessPurp2$Leaf.Number))

SN1 <- lmer(SeedNumber~Trt + Block + Leaf.Number + Block:Trt + (1|ML), FitnessPurp2)
anova(SN1) # Treatment is significant effect on plant seed number; Significant Treatment by Block effect

ggplot(FitnessPurp, aes(Trt, RelativeFit, fill=Block)) +
  geom_boxplot() +
  scale_fill_brewer("Paired") +
  theme_classic() +
  ylab("Relative Fitness") +
  ggtitle("Relative Fitness by Treatment and Block")

# Remove size effects from fitness
StdFitness <- FitnessPurp2[c("Trt", "Species", "Block", "ML", "RelativeFit", "Leaf.Number")] # Subset fitness data for variables of interest

# Run one-way ANOVA to remove size effect--i.e., keep residuals
StdFitness$RelativeFitness <- residuals(lm(RelativeFit ~ Leaf.Number, FitnessPurp2)) 

# Compare residuals and non-standard values of fitness
plot(StdFitness$RelativeFit, StdFitness$RelativeFitness)

# Average fitness by block, maternal line and treatment--we use NON standardized fitness (ie size effects not removed)
FitAveraged = aggregate(RelativeFit ~ Block + Trt + ML, FitnessPurp, mean)
colnames(FitAveraged) <- c("Block", "Trt", "ML", "RelativeFitness")
FitAveraged$TRT <- FitAveraged$Trt
dim(FitAveraged)
head(FitAveraged)

# Average size
SizeAveraged <- aggregate(Leaf.Number ~ Block + Trt + ML + Species, LeafData,mean)
colnames(SizeAveraged) <- c("Block", "Trt", "ML", "Species", "Size")

SizeAveraged$TRT <- SizeAveraged$Trt
dim(SizeAveraged)
head(SizeAveraged)

SizePurp <- SizeAveraged %>% filter(Species == "Ip")
```


```{r, fig.height = 12, fig.width = 16}

####  ####  ####  ####  ####  ####   ####
#  Examine selection on microbiome first
####  ####  ####  ####  ####  ####   ####
FitnessPurp$TRT <- FitnessPurp$Trt
FitnessPurp$Combos <- as.character(FitnessPurp$Combos)
FitnessPurp[which(FitnessPurp$TRT == "Alone"),]$Combos <- "none"
FitnessPurp$Combos <- as.factor(FitnessPurp$Combos)

BrayFit <- merge(physeq.pcoa.df,FitAveraged)

FitAlpha <- merge(FitAveraged, alpha)
dim(FitAlpha)

ggplot(FitAlpha, aes(TRT, RelativeFitness)) +
  geom_boxplot() +
  scale_fill_brewer("Paired") +
  theme_classic() +
  ylab("Relative Fitness") +
  ggtitle("Relative Fitness by Treatment and Block") +
  facet_grid(~Block)


# Combine with root data
library(dplyr)

RootAveraged <- aggregate(list(RootData[c("PC1", "PC2", "PC3", "PC4")]),by=list(RootData$Trt, RootData$ML),FUN=mean) 

colnames(RootAveraged) <- c("Trt", "ML", "PC1", "PC2", "PC3", "PC4")
head(RootAveraged)

RootFitAlpha <- merge(FitAlpha, RootAveraged)
head(RootFitAlpha)

#RootFitLeafAlpha=merge(RootFitAlpha, LeafData[c("Sample_ID", "Leaf.Number")],by="Sample_ID")

#RootFitLeafAlpha$Leaf.Number=as.numeric(as.character(RootFitLeafAlpha$Leaf.Number))


# CombinE root,fitness/bray estimates
RootFitBray <- merge(BrayFit, RootAveraged)


# Plot boxplots of averaged root traits
PC2_Box <- ggplot(RootAveraged, aes(x = "", y = PC2)) +
  geom_boxplot() +
  #geom_jitter() +
  xlab("Root Architecture") +
  theme_classic() +
  Tx+ 
  coord_flip()

# Plot boxplots of averaged root traits
PC1_Box=ggplot(RootAveraged, aes(x = "", y = PC1)) +
  geom_boxplot() +
  #geom_jitter() +
  xlab("Root Topology") +
  theme_classic() +
  Tx+ 
  coord_flip()

# Plot boxplots of averaged root traits
PC4_Box=ggplot(RootAveraged, aes(x = "", y = PC4)) +
  geom_boxplot() +
  #geom_jitter() +
  xlab("Root Morphology") +
  theme_classic() +
  Tx+ 
  coord_flip()

#ggarrange(P2.rich,P2.even,P4.Sim,P4.simIn,nrow=2,ncol=2)
AB <- cowplot::plot_grid(P2.rich,P2.even, align = "hv", ncol = 2, Labels = c("A", "B"), label_size = 22 ,hjust = -11,vjust = 1, Label_x = -0.16)

ab <- cowplot::plot_grid(P4.Sim,P4.simIn, align = "hv", ncol = 2, Labels =c("A", "B"), label_size =22 ,hjust = -11,vjust=1, Label_x = -0.17)


P4.simIn +
  xlab("Root morphology (PC4)")
# Common x title

x.grob <- textGrob("Root morphology (PC4)", 
                   gp=gpar(col="black", fontsize = 25), rot=0)

gridExtra::grid.arrange(gridExtra::arrangeGrob(ab,bottom = x.grob, padding = unit(0.5,units = 'in'), nrow=1))

# Common x title
x.grob1 <- textGrob("Root architecture (PC2)", 
                   gp = gpar(col="black", fontsize = 25), rot = 0)

gridExtra::grid.arrange(gridExtra::arrangeGrob(AB, bottom = x.grob1, padding = unit(0.5,units = 'in'), nrow=1))

```


## Selection on microbe variables

```{r}
# Quadratic microbe term
FitAlpha$even2 <- (FitAlpha$even*FitAlpha$even)
FitAlpha$rich2 <- (FitAlpha$rich*FitAlpha$rich)
FitAlpha$InSim2 <- (FitAlpha$InvSimp*FitAlpha$InvSimp)
FitAlpha$sim2 <- (FitAlpha$sim*FitAlpha$sim)

```

### Selection on richness
```{r cor on richness}
summary(lm(RelativeFitness ~ Block +rich+rich2, FitAlpha %>% filter(TRT == "Alone")))
summary(lm(RelativeFitness ~ Block +rich+rich2, FitAlpha %>% filter(TRT != "Alone")))

richness_sel <- lm(RelativeFitness ~ TRT*Block +rich, FitAlpha)
summary(richness_sel) # No evidence of indirect selection
anova(richness_sel)


ggplot(FitAlpha) +
  geom_point(aes(RelativeFitness, rich), alpha = 0.5, size = 5) +
  geom_smooth(method = "lm", aes(RelativeFitness, rich), fullrange = TRUE) +
  theme_classic() +
  ylab("Relative fitness") +
  xlab("Richness") +
  Tx2

```


### Selection on Inverse Simpson
```{r cor on Inverse Simpson}
summary(lm(RelativeFitness ~ Block +InvSimp+InSim2, FitAlpha %>% filter(TRT == "Alone")))

summary(lm(RelativeFitness ~ Block +InvSimp+InSim2, FitAlpha %>% filter(TRT != "Alone")))

Invn_sel <- lm(RelativeFitness ~ TRT*Block + InvSimp + Block, FitAlpha)
summary(Invn_sel) # No evidence of indirect selection

ggplot(FitAlpha) +
  geom_point(aes(RelativeFitness, InvSimp), alpha = 0.5, size = 5) +
  geom_smooth(method = "lm", aes(RelativeFitness, InvSimp), fullrange = TRUE) +
  theme_classic() +
  ylab("Relative fitness") +
  xlab("Inverse Simpson") +
  Tx2
```

## Selection on Simpson
```{r cor on Simpson }
summary(lm(RelativeFitness ~ Block + sim + sim2, FitAlpha %>% filter(TRT == "Alone")))

summary(lm(RelativeFitness ~ Block + sim + sim2, FitAlpha %>% filter(TRT != "Alone")))

sim_sel <- lm(RelativeFitness ~ TRT*Block + sim, FitAlpha)
summary(sim_sel) # NS

ggplot(FitAlpha) +
  geom_point(aes(RelativeFitness, sim), alpha = 0.5, size = 5) +
  geom_smooth(method = "lm", aes(RelativeFitness, sim), fullrange = TRUE) +
  theme_classic() +
  ylab("Relative fitness") +
  xlab("Simpson") +
  Tx2
```

## Selection on Evenness
```{r cor on Evenness}

summary(lm(RelativeFitness ~ Block + even + even2 , FitAlpha %>% filter(TRT == "Alone")))

summary(lm(RelativeFitness ~ Block + even + even2, FitAlpha %>% filter(TRT != "Alone"))) # Significant quadratic term for evenness on fitnes (super low slope tho)

even_sel <- lm(RelativeFitness ~ TRT*Block + TRT*even2 + even, FitAlpha) # if we include block by treatment interaction there is marginal significance for a treatment by evenness squared effect
even_sel2 <- lm(RelativeFitness ~ Block + TRT*even2 + even, FitAlpha) # if we include block by treatment interaction there is marginal significance for a treatment by evenness squared effect

anova(even_sel, even_sel2) # Based on RSS--second model/simpler one is a bit better
anova(even_sel) # NS

ggplot(FitAlpha) +
  geom_point(aes(RelativeFitness, even), alpha = 0.5, size = 5) +
  stat_smooth(method = "lm",formula = y ~poly(x, 2), aes(RelativeFitness,even)) +
  theme_classic() +
  ylab("Relative fitness") +
  xlab("Evenness") +
  Tx2
```

```{r}

# Remove block effects from fitness
RootFitAlpha$FitResidBlk <- resid(lm(RelativeFitness ~ Block, RootFitAlpha))

RootFitAlpha$Comp <- sub(".*\\-", "", RootFitAlpha$Combos)

```

# ANCOVA

```{r}


#RootFitAlpha<-merge(SizePurp, RootFitAlpha)

library(interactions)
library(jtools)

# Scale the microbial variables
RootFitAlpha$richScaled<-scale(RootFitAlpha$rich)
RootFitAlpha$SimScaled<-scale(RootFitAlpha$sim)
RootFitAlpha$InvSimScaled<-scale(RootFitAlpha$InvSimp)
RootFitAlpha$EvenScaled<-scale(RootFitAlpha$even)


# ANCOVA MUltivariate Linear Regressions


ANCOVA<-(lm(RelativeFitness~TRT + Block+ TRT:Block + PC1*TRT+PC2*TRT+PC3*TRT+PC4*TRT+richScaled*TRT+InvSimScaled*TRT+EvenScaled*TRT+PC1*Block + PC2*Block + PC3*Block + PC4*Block+richScaled*Block+InvSimScaled*Block+EvenScaled*Block, RootFitAlpha)) # Full model reported

# Step wise backward regression
library(MASS)

step<-stepAIC(ANCOVA,direction = "backward",trace=FALSE)
step$anova



model<-lm(RelativeFitness ~ TRT + Block + PC1 + PC2 + PC3 + PC4 + richScaled + 
    InvSimScaled + EvenScaled + TRT:Block + TRT:PC1 + TRT:PC3 + 
    TRT:richScaled + TRT:EvenScaled + Block:PC1 + Block:PC2 + 
    Block:PC3 + Block:PC4 + Block:richScaled + Block:InvSimScaled + 
    Block:EvenScaled, RootFitAlpha
)


# anova(model) # Type I sum of squres--sequence order matters, interactions not accounted

car::Anova(model, type="III") # Report the type three sums of squares~
summary(model)
# summary(ANCOVA)




summary(lm(RelativeFitness~rich, RootFitAlpha %>% filter(TRT == "Alone"))) # Alone
summary(lm(RelativeFitness~rich, RootFitAlpha %>% filter(TRT != "Alone"))) # Competition


# Plot
EvenPlot1<-ggplot() +
  geom_point(data=RootFitAlpha %>% filter(TRT == "Alone"), aes(EvenScaled,RelativeFitness), size = 3, alpha=0.5) +
#    geom_smooth(data=RootFitAlpha %>% filter(TRT == "Alone"), aes(even,RelativeFitness),method = "lm", fullrange = TRUE, se=FALSE) +
  geom_abline(slope=-0.47, intercept = 1.18,color="blue", size = 2) +
  theme_classic() +
  xlab("Evenness") +
  Tx+
  theme(axis.text.x = element_text(angle=45), axis.text = element_text(color="black", size=15,vjust = 0.5,hjust=1)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5))


EvenPlot2<-ggplot() +
  geom_point(data=RootFitAlpha %>% filter(TRT=="Inter"), aes(EvenScaled,RelativeFitness), size = 3, alpha=0.5) +
#    geom_smooth(data=RootFitAlpha %>% filter(TRT=="Inter"), aes(EvenScaled,RelativeFitness),method = "lm", fullrange = TRUE, se=FALSE) +
    geom_abline(slope=-0.28, intercept = 1.03,color="blue", size = 2) +

  theme_classic() +
  xlab("Evenness") +
  Tx+
  theme(axis.text.x = element_text(angle=45), axis.text = element_text(color="black", size=15,vjust = 0.5,hjust=1)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5))

ggplot() +
  geom_point(data=RootFitAlpha, aes(EvenScaled,RelativeFitness,color=TRT), size = 3, alpha=0.5) +
#    geom_smooth(data=RootFitAlpha %>% filter(TRT=="Inter"), aes(EvenScaled,RelativeFitness),method = "lm", fullrange = TRUE, se=FALSE) +
    geom_abline(slope=-0.47, intercept = 1.18,color="red", size = 2, Linetype="dashed") +

    geom_abline(slope=-0.28, intercept = 1.03,color="black", size = 2) +

  theme_classic() +
  xlab("Evenness") +
  Tx2+
  theme(axis.text.x = element_text(angle=45), axis.text = element_text(color="black", size=15,vjust = 0.5,hjust=1)) +
  scale_x_continuous(breaks = scales::pretty_breaks(n = 5)) + scale_color_manual(values = c("red", "black"), "Treatment", labels = c("Alone", "Competition"))


#ggarrange(P2.rich,P2.even,P4.Sim,P4.simIn,nrow=2,ncol=2)
cowplot::plot_grid(EvenPlot1,EvenPlot2, align = "hv", ncol = 2, Labels =c("A", "B"), label_size =22 ,hjust = -9,vjust=2, Label_x = -0.16)




Block1<-ggplot() +
  geom_point(data=RootFitAlpha %>% filter(Block=="1"), aes(even,RelativeFitness), size = 3, alpha=0.5) +
    geom_smooth(data=RootFitAlpha %>% filter(Block=="1"), aes(even,RelativeFitness), se=FALSE,method = "lm", fullrange = TRUE) +
    theme_classic() +
  xlab("Eveness") +
  Tx+
  theme(axis.text.x = element_text(angle=45), axis.text = element_text(color="black", size=15,vjust = 0.5,hjust=1))
   
Block2<-ggplot() +
  geom_point(data=RootFitAlpha %>% filter(Block=="2"), aes(even,RelativeFitness), size = 3, alpha=0.5) +
    geom_smooth(data=RootFitAlpha %>% filter(Block=="2"), aes(even,RelativeFitness), se=FALSE,method = "lm", fullrange = TRUE) +
    theme_classic() +
  xlab("Eveness") +
  Tx+
  theme(axis.text.x = element_text(angle=45), axis.text = element_text(color="black", size=15,vjust = 0.5,hjust=1))


Block3<-ggplot() +
  geom_point(data=RootFitAlpha %>% filter(Block=="3"), aes(even,RelativeFitness), size = 3, alpha=0.5) +
    geom_smooth(data=RootFitAlpha %>% filter(Block=="3"), aes(even,RelativeFitness), se=FALSE,method = "lm", fullrange = TRUE) +
    theme_classic() +
  xlab("Eveness") +
  Tx+
  theme(axis.text.x = element_text(angle=45), axis.text = element_text(color="black", size=15,vjust = 0.5,hjust=1))


Block4<-ggplot() +
  geom_point(data=RootFitAlpha %>% filter(Block=="4"), aes(even,RelativeFitness), size = 3, alpha=0.5) +
    geom_smooth(data=RootFitAlpha %>% filter(Block=="4"), aes(even,RelativeFitness),method = "lm", fullrange = TRUE) +
    theme_classic() +
  xlab("Eveness") +
  Tx+
  theme(axis.text.x = element_text(angle=45), axis.text = element_text(color="black", size=15,vjust = 0.5,hjust=1))

```






# MANTEL  (Table 4)
```{r}  

#  Here we use the Family mean values of root traits---this would indicate evidence for 'phenotypic selection' 

  # Load matrix Bray distances of OTU abundance
  OTU_table = t(otu_table(physeq1)) # Write out OTU table
  

  # Pull out Root traits of interest and save these to the Sampled data frame
  SampledFit = merge(FitAveraged, sampledf, by = c("TRT", "ML", "Block"))


# Pull out Root traits of interest and save these to the Sampled data frame
SampledRoots = merge(RootAlphaObs, sampledf)
SampledRootsIp = subset(SampledRoots, Species == "Ip")

OTU_table = OTU_table[row.names(OTU_table) %in% SampledRootsIp$Sample_ID,]


    OTU = OTU_table
    Roots = SampledRootsIp

  # Calculate root architecture distances with euclidean distance 

    PC2 = SampledRootsIp$PC2 # isolate PC2, i.e. root architecture
    PC2.dist = dist(PC2)

  # Calculate Bray distance matrix for OTU table

    Bray = vegdist(OTU, method = "bray")


    
# OTU Bray vs Root architecture

    OTU_pc2 = mantel(Bray, PC2.dist, method = "spearman", permutations = 9999, na.rm = TRUE)

    # Marginally significant--very low r value
    # Mantel statistic r: 0.06836  
    # Significance: 0.07
    
# Repeat for root topology, size and morphology

  PC1 = SampledRootsIp$PC1 # isolate 
  PC1.dist = dist(PC1)
  PC3 = SampledRootsIp$PC3 # isolate 
  PC3.dist=dist(PC3)
  PC4 = SampledRootsIp$PC1 # isolate 
  PC4.dist = dist(PC4)

# Examine overall distances in root system; i.e. use all PCs in distance calculation

    PCall = SampledRootsIp[grep("PC",names(SampledRootsIp))]
    PC.dist = dist(PCall)

    OTU_pc = mantel(Bray, PC.dist, method = "spearman", permutations = 9999, na.rm = TRUE)

###             examine mantel test using observed values of roots ###

OTU_pc1 = mantel(Bray, PC1.dist, method = "spearman", permutations = 9999, na.rm = TRUE)
# OTU_pc2 = mantel(Bray,PC2.dist, method = "spearman", permutations = 9999, na.rm = TRUE)
OTU_pc3 = mantel(Bray, PC3.dist, method = "spearman", permutations = 9999, na.rm = TRUE)
OTU_pc4 = mantel(Bray, PC4.dist, method = "spearman", permutations = 9999, na.rm = TRUE)
```

PC1 Mantel

```{r}
OTU_pc1
```

PC2 Mantel

```{r}
OTU_pc2
```


PC3 Mantel

```{r}
OTU_pc3
```


PC4 Mantel

```{r}
OTU_pc4
```



